---
title: "Analysis_Hlil"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Programs and working directory

```{r, message=FALSE}
##
#install.packages("adegenet", repos = universe)
library(adegenet)
##
library(tidyverse)
library(vcfR)
library(SNPfiltR)
#install.packages("MasterBayes")
#library(apparent)
#install.packages("gtools")
library(gtools)
```

## Convert the vcf to objects useful downstream

```{r}
vcfR <- read.vcfR("/storage1/fs1/christine.e.edwards/Active/mashburn/Hibiscus/3_Hlil/stacks/rerun/Hlil_rerun.vcf.gz")
Hlil.genind <- vcfR2genind(vcfR)
Hlil.genclone <- poppr::as.genclone(Hlil.genind)
Hlil.genlight <- vcfR2genlight(vcfR)
```

## PCA using snapclust

```{r}
# create the popmap
popmap<-data.frame(id=colnames(vcfR@gt)[2:length(colnames(vcfR@gt))],pop=substr(colnames(vcfR@gt)[2:length(colnames(vcfR@gt))], 3,11))
```

Transform the popmap with actual pop data

```{r}
# make a copy popmap
popmap1 = popmap
# make a list with sample names to help with pop assignment
sample.list <- as.list(popmap1[1])

# make lists for each population
outplanted <- c("Hlil130", "Hlil132", "Hlil137", "Hlil143", "Hlil145", "Hlil146", "Hlil147", "Hlil148", "Hlil150", "Hlil152", "Hlil153", "Hlil154", "Hlil155", "Hlil156", "Hlil157", "Hlil158", "Hlil159", "Hlil160", "Hlil161", "Hlil162", "Hlil163", "Hlil164", "Hlil165", "Hlil166", "Hlil179", "Hlil180", "Hlil181", "Hlil182", "Hlil183", "Hlil184", "Hlil185", "Hlil186", "Hlil188", "Hlil189", "Hlil190", "Hlil191", "Hlil193", "Hlil195", "Hlil196", "Hlil197", "Hlil198", "Hlil199", "Hlil200", "Hlil201", "Hlil202", "Hlil203", "Hlil204", "Hlil205", "Hlil207", "Hlil208", "Hlil209", "Hlil210", "Hlil211", "Hlil212", "Hlil213", "Hlil214", "Hlil215", "Hlil216", "Hlil217", "Hlil219", "Hlil220", "Hlil221", "Hlil224", "Hlil225", "Hlil226", "Hlil228", "Hlil229", "Hlil231", "Hlil232", "Hlil233", "Hlil235", "Hlil239", "Hlil240", "Hlil241", "Hlil242", "Hlil243", "Hlil244", "Hlil245", "Hlil246", "Hlil247", "Hlil248")
clones <- c("Hlil167", "Hlil168", "Hlil169", "Hlil170", "Hlil171", "Hlil172", "Hlil173", "Hlil174", "Hlil175", "Hlil176", "Hlil177", "Hlil178")
ex_situ <- c("HlilNBG351", "HlilNTBG260", "HlilWBG226")
wild <- c("HlilW222", "HlilW223")

# the worst for loop in the world to add pop names
for (i in 1:length(popmap1$id)) {
  sample <- popmap1$id[i]
  if (sample %in% outplanted) {
    pop <- "outplanted"
  } else if (sample %in% clones) {
    pop <- "clones"
  } else if (sample %in% ex_situ) {
    pop <- "ex-situ"
  } else if (sample %in% wild) {
    pop <- "wild"
  }
  popmap1$pop[i] <- pop
}

```


```{r}
# pca using SNPfiltR
assess_missing_data_pca(vcfR, popmap = popmap1, clustering = FALSE)
```

PCA looking at genetic groups

```{r}
# make a copy popmap
popmap2 = popmap
# make a list with sample names to help with pop assignment
sample.list <- as.list(popmap2[1])

# make lists for each population
cross_1x2 <- c("Hlil130", "Hlil132", "Hlil145", "Hlil146", "Hlil147", "Hlil150", "Hlil154", "Hlil155", "Hlil160", "Hlil165", "Hlil182", "Hlil214", "Hlil248")
cross_1x3 <- c("Hlil137", "Hlil157", "Hlil162", "Hlil163", "Hlil164", "Hlil166", "Hlil179", "Hlil180", "Hlil183",  "Hlil185", "Hlil186",  "Hlil190", "Hlil191", "Hlil195", "Hlil196", "Hlil197", "Hlil198", "Hlil199", "Hlil200", "Hlil202", "Hlil203", "Hlil204", "Hlil205", "Hlil207", "Hlil208", "Hlil209", "Hlil210", "Hlil213", "Hlil215", "Hlil216", "Hlil217", "Hlil220", "Hlil221", "Hlil224", "Hlil225", "Hlil226", "Hlil228", "Hlil229",  "Hlil232", "Hlil233", "Hlil240", "Hlil241", "Hlil242")
cross_2x3 <- c("Hlil184", "Hlil188", "Hlil189", "Hlil193", "Hlil231", "Hlil235", "Hlil239", "Hlil245", "Hlil247")
self_1x1 <- c("Hlil143", "Hlil148", "Hlil152", "Hlil153", "Hlil156", "Hlil158", "Hlil159", "Hlil161", "Hlil211", "Hlil243", "Hlil244", "Hlil246")
self_2x2 <- c("Hlil181", "Hlil212", "Hlil219")
clone_group_1 <- c("Hlil167", "Hlil170", "Hlil171", "Hlil172", "Hlil176", "Hlil177", "Hlil178", "HlilNTBG260", "HlilWBG226")
clone_group_2 <- c("Hlil173", "Hlil174", "Hlil175", "HlilW223", "HlilNBG351")
clone_group_3 <- c("Hlil168", "Hlil169")
clone_group_4 <- c("HlilW222", "Hlil201")

# the worst for loop in the world to add pop names
for (i in 1:length(popmap$id)) {
  sample <- popmap2$id[i]
  if (sample %in% cross_1x2) {
    pop <- "Cross 1x2"
  } else if (sample %in% cross_1x3) {
    pop <- "Cross 1x3"
  } else if (sample %in% cross_2x3) {
    pop <- "Cross 2x3"
  } else if (sample %in% self_1x1) {
    pop <- "Self 1x1"
  } else if (sample %in% self_2x2) {
    pop <- "Self 2x2"
  } else if (sample %in% clone_group_1) {
    pop <- "Genotype Group 1"
  } else if (sample %in% clone_group_2) {
    pop <- "Genotype Group 2"
  } else if (sample %in% clone_group_3) {
    pop <- "Genotype Group 3"
  } else if (sample %in% clone_group_4) {
    pop <- "Genotype Group 4"
  }
  popmap2$pop[i] <- pop
}

```


```{r}
# pca using SNPfiltR
assess_missing_data_pca(vcfR, popmap = popmap2, clustering = FALSE)
```

## PCA with plink

```{r}
# read in data for the pca done by plink
pca <- read_table2("/storage1/fs1/christine.e.edwards/Active/mashburn/Hibiscus/3_Hlil/pca/Hlil_rerun.eigenvec", col_names = FALSE)
eigenval <- scan("/storage1/fs1/christine.e.edwards/Active/mashburn/Hibiscus/3_Hlil/pca/Hlil_rerun.eigenval")
```

```{r}
## clean up the pca tibble
# remove the repeat sample name column
pca <- pca[,-1]
# set the first column to be called 'ind'
names(pca)[1] <- "ind"
# set the other column names to include 'PC'
names(pca)[2:ncol(pca)] <- paste0("PC", 1:(ncol(pca)-1))
# later: add population data
```


```{r}
## Plot variance explained
# translate eigenvalues into percentage variance explained
pve <- data.frame(PC = 1:20, pve = eigenval/sum(eigenval)*100)
# make a bar plot
a <- ggplot(pve, aes(PC, pve)) + geom_bar(stat = "identity")
a + ylab("Percentage Variance Explained") + theme_light()
# calculate the cumulative sum of pve
cumsum(pve$pve)
#> it's 100% !
```


```{r}
# Plot the PCA
b <- ggplot(pca, aes(PC1, PC2)) + geom_point(size = 3) + coord_equal() + theme_light()
b + xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)"))
```

## DAPC

https://github.com/thibautjombart/adegenet/wiki

Use DAPC to infer the number of clusters of genetic relatedness. Data is transformed using PCA and then clusters are identified with discriminant analysis (DA). DAPC is similar to STRUCTURE but better if there are clonal individuals/populations.

```{r, message=FALSE}
# identify ideal number of clusters using BIC score
grp.bic <- find.clusters(Hlil_genlight, n.pca = 100, choose = FALSE, stat = "BIC")
plot(grp.bic$Kstat, type = "o", xlab = "number of clusters (K)", ylab = "BIC", main = "find.clusters on genlight object\n(two groups)")

# identify ideal number of clusters using AIC score
grp.aic <- find.clusters(Hlil_genlight, n.pca = 100, choose = FALSE, stat = "AIC")
plot(grp.aic$Kstat, type = "o", xlab = "number of clusters (K)", ylab = "AIC", main = "find.clusters on genlight object\n(two groups)")

#pop(Hlil_genind)
#dapc.Hlil <- dapc(Hlil_genind, pop = NULL, var.contrib = TRUE, scale = FALSE, n.pca = NULL, n.da = NULL)
#scatter(dapc.Hlil, cell = 0, pch = 18:23, cstar = 0, mstree = TRUE, lwd = 2, lty = 2)
```




```{r}
# make a list with sample names
sample.list <- scan("/storage1/fs1/christine.e.edwards/Active/mashburn/Hibiscus/3_Hlil/introgress/samples.txt", character(), quote = "")

# make lists for each population
outplanted <- c("Hlil130", "Hlil132", "Hlil135", "Hlil137", "Hlil143", "Hlil145", "Hlil146", "Hlil147", "Hlil148", "Hlil150", "Hlil152", "Hlil153", "Hlil154", "Hlil155", "Hlil156", "Hlil157", "Hlil158", "Hlil159", "Hlil160", "Hlil161", "Hlil162", "Hlil163", "Hlil164", "Hlil165", "Hlil166", "Hlil179", "Hlil180", "Hlil181", "Hlil182", "Hlil183", "Hlil184", "Hlil185", "Hlil186", "Hlil188", "Hlil189", "Hlil190", "Hlil191", "Hlil193", "Hlil195", "Hlil196", "Hlil197", "Hlil198", "Hlil199", "Hlil200", "Hlil201", "Hlil202", "Hlil203", "Hlil204", "Hlil205", "Hlil206", "Hlil207", "Hlil208", "Hlil209", "Hlil210", "Hlil211", "Hlil212", "Hlil213", "Hlil214", "Hlil215", "Hlil216", "Hlil217", "Hlil218", "Hlil219", "Hlil220", "Hlil221", "Hlil224", "Hlil225", "Hlil226", "Hlil227", "Hlil228", "Hlil229", "Hlil230", "Hlil231", "Hlil232", "Hlil233", "Hlil235", "Hlil236", "Hlil237", "Hlil238", "Hlil239", "Hlil240", "Hlil241", "Hlil242", "Hlil243", "Hlil244", "Hlil245", "Hlil246", "Hlil247", "Hlil248")
pot_clones <- c("Hlil167", "Hlil168", "Hlil169", "Hlil170", "Hlil171", "Hlil172", "Hlil173", "Hlil174", "Hlil175", "Hlil176", "Hlil177", "Hlil178")
ex_situ <- c("HlilNBG351", "HlilNTBG260", "HlilWBG226")
wild <- c("HlilW222", "HlilW223")

# the worst for loop in the world to add pop names
popmap <- as.list(rep(NA, length(sample.list)))
for (i in 1:length(sample.list)) {
  sample <- sample.list[i]
  if (sample %in% outplanted) {
    pop <- "outplanted"
  } else if (sample %in% pot_clones) {
    pop <- "pot_clones"
  } else if (sample %in% ex_situ) {
    pop <- "ex_situ"
  } else if (sample %in% wild) {
    pop <- "wild"
  }
  popmap[[i]] <- c(sample, pop)
}

## i now have an object (list of vectors) that i can use to fill in the genotypes object for introgress
## see the example data 2 from http://www.uwyo.edu/buerkle/software/introgress/
```

## GENHET

Upload data in the appropriate format for GENHET
```{r}
# import genhet object
Hlil.genhet <- read.table("/storage1/fs1/christine.e.edwards/Active/mashburn/Hibiscus/3_Hlil/genhet/Hlil.genhet", sep = "\t", header = T)

# use genhet object to create a vector of locus names
# unneccessarily complicated
odd <- seq_len(ncol(Hlil.genhet[-1])) %% 2
locus.names <- gsub("a", "", names(Hlil.genhet[-1])[odd == 1])

```

Perform GENHET
```{r}
genhet.matrix <- as.matrix(GENHET(dat = Hlil.genhet, estimfreq = "T", locname = locus.names))
# save the output
write.table(genhet.matrix, "/storage1/fs1/christine.e.edwards/Active/mashburn/Hibiscus/3_Hlil/genhet/Hlil_genhet.csv")
```

